---
title: "04-Enrichment"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("enrichR")
#install.packages("openxlsx")# independant of java
#install.packages("devtools") 
#devtools::install_github("wjawaid/enrichR")
library(enrichR)
library(openxlsx)

# NOTE - if you have an error "EnrichR website not responding"  and unable to connect to EnrichR database online, try this
# OPTION 1
#devtools::install_github("wjawaid/enrichR")
#library(enrichR)

# OPTION 2 - update curl and wget
# (Step a) Open the "terminal" application for these commands. Follow instructions for mac
# https://help.ubidots.com/en/articles/2165289-learn-how-to-install-run-curl-on-windows-macosx-linux
# 
# (Step b) Install/update wget to the latest version
# Use the same "terminal" application. Enter
# brew install wget
# 
# (step C) Restart your computer. 
# Try out these EnrichR commands to see if you are able to access the enrichR databases online without error message
# install.packages("enrichR")
# library(enrichR)
# enrichR::listEnrichrDbs()

```

## R Markdown - Enrichment anlalysis using EnrichR R package

Summary

-   I have already shortlisted the genes in my
    Tazeen_GroupCompAnalysis.Rmd file based on p-value cut off of 0.01

-    I also have created a clean list of gene names in the same file
    above

-   However, I will be redoing the filtering and gene name clean up to
    ensure I get the same results, and also to make sure the gene names
    are given to the EnrichR function in the correct/expected format

-   I won't be writing additional files though for step 1 and 2 as the
    outputs are already there and I just need to check the content and
    formats

-   Then I will connect to the EnrichR database and run Enrichment

## Import T-test results so that we can short list further

```{r}
#folder that contains group comparison results
fileName <- "input/Tazeen_Moderate_Incipient_Ttest_Shortlisted.csv"

# read in group comparison results
result1 <- read.csv(file = fileName, 
                        header = T, stringsAsFactors = F, row.names = 1)
```

## Filtering/shortlisting Short list results based in p-value cut off (pvalue \<= 0.01) This is already done but redoing just to be sure

```{r}
pValueCutOff <- 0.01
which1 <- ((as.numeric(result1$Pvalue) <= pValueCutOff))
table(which1) #146 genes with p <= 0.01. 

resultShort <- result1[which1, ] #short listed genes 
```

## Clean gene names

Note - the gene names are in the form of "reporterid_genename". So need
to split this.

```{r}
funcSplit <- function(rep_gene) {
    rep_gene_split <- unlist(strsplit(x = rep_gene, 
                                      split = "|", 
                                      fixed = TRUE))
    gene <- rep_gene_split[2]
    return(gene)
}
geneListSplit <- apply(X = as.matrix((resultShort$Feature)), 
                       MARGIN = 1, FUN = funcSplit )
head(geneListSplit) #cleaned gene names

#remove duplicates
geneListSplit1 <- unique(geneListSplit)

# remove NA value
geneListFinal <- na.omit(geneListSplit1) #156 genes

head(geneListFinal)
```

```{r}
#print number of unique genes
length(geneListFinal)
```

## Load Databases for Enrichr R package , and check connection

```{r}
#checking if EnrichR website and packing are working
#dbs <- enrichR::listEnrichrDbs()  #total number of databases available = 200+ 

#testing if EnrichR package is working
testOutput <- enrichR::enrichr(genes = c("Runx1", "Gfi1", "Gfi1b", "Spi1", "Gata1", "Kdr"), databases = "KEGG_2021_Human")
head(testOutput[[1]])

#List of databases for which enrichment analysis will be run
dblist1 <- read.csv(file = "input/2023-EnrichR-Databases.txt", 
                    header = F, stringsAsFactors = F)

head(dblist1)
```

## Call function to run Enrichment

```{r, message=FALSE, warning=FALSE}
# set output file name
outputFileName1 <- paste("output/Tazeen_Moderate_Incipient", "_EnrichR.xlsx", sep="")

#Load R script into the environment
source(file = "functionEnrichment.R") 

#call function to run Enrichment
functionEnrichment(dblist1, geneListFinal, outputFileName1)

#NEED INTERNET CONNECTION
```

Note - you will need internet connection to complete the above step.
